---
title: "Visualization Examples (gstat backend)"
author: "James Otto"
date: "6/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      results = "hide")
```

## Simulated Examples
### Constant Mean (Ordinary Kriging)

First, we simulate data from a GRF with mean 0 and an exponential covariance structure with parameters $\sigma^2 = 5$ and $\phi = 30$.
We simulate observations on a sample of size 600 from the set $D = \{1, \ldots, 100\} \times \{1, \ldots, 100\} \subset \mathbb{R}^2$.
```{r simulating_OK}
library("tidyverse"); theme_set(theme_bw())
library("patchwork")

set.seed(2)

grid <- expand.grid(1:100, 1:100) %>% 
  tibble() %>% 
  rename(x = Var1, y = Var2)

dom <- slice_sample(grid, n = 600)

data <- geoR::grf(grid = dom, cov.model = "exp", cov.pars = c(5, 30), nugget = 1)$data %>%
  {tibble(z = .)}

df <- bind_cols(dom, data)
```

Below, we have visualized the simulated data and plotted the empirical variogram.
```{r visualizing_sim_OK}
ggplot(df, aes(x, y, color = z, size = z)) +
  geom_point() +
  scale_color_viridis_c() +
  guides(color = guide_legend(), size = guide_legend())

vgm <- gstat::variogram(z ~ 1, locations = ~x+y, data = df)
plot(vgm)
```

We can estimate parameters of the variogram and perform ordinary kriging to interpolate between the original data.
Note, we have to provide initial values for the parameters to be estimated as well as specify the exponential covariance model.
```{r manual_kriging}
vgm_fit <- gstat::fit.variogram(vgm, model = gstat::vgm(1, "Exp", 5, 30))

kriged <- gstat::krige(formula = z ~ 1, data = df, locations = ~x+y, newdata = grid, model = vgm_fit)
names(kriged) <- c("x", "y", "prediction", "var")

ggplot(df, aes(x, y, size = z)) +
  geom_raster(data = kriged, aes(x, y, fill = prediction), inherit.aes = FALSE) +
  geom_point() +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE) +
  scale_fill_viridis_c() +
  guides(size = guide_legend(reverse = TRUE))
```

***

Now, we load in the functions which extend `ggplot2`:
```{r loading_functions}
source(here::here("R/geom-krige.R"))
source(here::here("R/stat-krige.R"))
source(here::here("R/geom-krige-contour.R"))
source(here::here("R/stat-krige-contour.R"))
```

Using these functions, we can create the same visualizations much easier.
Below, see that the kriging is handled by `geom_krige` -- we only need supply the model specification and initial values for the parameters.
Recall, these can be determined by visualizing the empirical variogram as before.
```{r geom_krige_OK_1}
ggplot(df, aes(x, y)) +
  geom_krige(aes(z = z), inits = c(6, 40, 1), model = "Exp") +
  geom_point(aes(size = z)) +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE) +
  scale_fill_viridis_c() +
  guides(size = guide_legend(reverse = TRUE))
```

We can adjust the resolution of the kriging as shown below:
```{r geom_krige_OK_2, fig.height = 10}
p_low <- ggplot(df, aes(x, y, z = z)) +
  geom_krige(nx = 50, ny = 50, inits = c(6, 40, 1), model = "Exp") +
  scale_fill_viridis_c() +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE) 

p_high <- ggplot(df, aes(x, y, z = z)) +
  geom_krige(nx = 200, ny = 200, inits = c(6, 40, 1), model = "Exp") +
  scale_fill_viridis_c() +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE)

p_low / p_high
```

We can also visualize the kriging output via contours with `geom_krige_contour` and `geom_krige_contour_filled` (which function similarly to `geom_contour` and `geom_contour_filled`):

```{r geom_krige_contour_OK_1, fig.width = 10}
p_contour <- ggplot(df, aes(x, y, z = z)) +
  geom_krige_contour(inits = c(6, 40, 1), model = "Exp") +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE)

p_contour_filled <- ggplot(df, aes(x, y, z = z)) +
  geom_krige_contour_filled(inits = c(6, 40, 1), model = "Exp") +
  scale_fill_viridis_d() +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE)

p_contour | p_contour_filled
```

We have also allowed for control of the resolution of the kriging here:
```{r geom_krige_contour_OK_2, fig.width = 10}
p_contour_lower <- ggplot(df, aes(x, y, z = z)) +
  geom_krige_contour(inits = c(6, 40, 1), model = "Exp", nx = 30, ny = 30, bins = 5) +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE)

p_contour_filled_lower <- ggplot(df, aes(x, y, z = z)) +
  geom_krige_contour_filled(inits = c(6, 40, 1), model = "Exp", nx = 30, ny = 30, bins = 5) +
  scale_fill_viridis_d() +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE)

p_contour_lower | p_contour_filled_lower
```


### Non-Constant Mean (Universal Kriging)

The extensions also allow for the mean of the GRF to depend on the coordinates.
Below, we simulate from such a process using the same set $D$ as before.
However, this time we are instead simulating from a process with Gaussian covariance and different covariance parameters.
```{r simulating_UK}
set.seed(1)
data <- geoR::grf(grid = dom, cov.model = "gau", cov.pars = c(10, 20), nugget = .5)$data %>%
  {tibble(z = .)}

df <- bind_cols(dom, data) %>% 
  mutate(z = z + 3/20 * x - 1/20 * y)
```

Below, we can see the data is spatially corellated in addition to the obvious linear trend:
```{r visualizing_sim_UK}
ggplot(df, aes(x, y, color = z, size = z)) +
  geom_point() +
  scale_color_viridis_c() +
  guides(color = guide_legend(), size = guide_legend())
```

In order to allow for this nonconstant mean, we perform Universal Kriging:

```{r geom_krige_UK_1}
ggplot(df, aes(x, y)) +
  geom_krige(aes(z = z), formula = z ~ x + y, inits = c(8, 25, 1), model = "Gau") +
  geom_point(aes(size = z)) +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE) +
  scale_fill_viridis_c() +
  guides(size = guide_legend(reverse = TRUE))
```

Below, we perform ordinary kriging and ignore the spatial structure of the mean.
See that the resulting interpolation does not fit the data as well, missing out on local variation.
```{r geom_krige_OK_incorrect}
ggplot(df, aes(x, y)) +
  geom_krige(aes(z = z), inits = c(8, 25, 1), model = "Gau") +
  geom_point(aes(size = z)) +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE) +
  scale_fill_viridis_c() +
  guides(size = guide_legend(reverse = TRUE))
```

***

Below, we show each of the different methods of visualizing universal kriging:
```{r geom_krige_UK_2, fig.height = 15}
p_heatmap_UK <- ggplot(df, aes(x, y)) +
  geom_krige(aes(z = z), formula = z ~ x + y, inits = c(8, 25, 1), model = "Gau", nx = 200, ny = 200) +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE) +
  scale_fill_viridis_c() 

p_contours_UK <- ggplot(df, aes(x, y)) +
  geom_krige_contour(aes(z = z), formula = z ~ x + y, inits = c(8, 25, 1), model = "Gau", nx = 30, ny = 30) +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE) 

p_filled_contours_UK <- ggplot(df, aes(x, y)) +
  geom_krige_contour_filled(aes(z = z), formula = z ~ x + y, inits = c(8, 25, 1), model = "Gau", nx = 30, ny = 30) +
  coord_cartesian(xlim = c(1, 100), ylim = c(1, 100), expand = FALSE) +
  scale_fill_viridis_d()

p_heatmap_UK /
p_contours_UK /
p_filled_contours_UK
```


### Latitude Longitude Data and ggmap

These extensions work well with the `ggmap` function.
Below, we simulate data in the Waco area:

```{r simulating_Waco}
set.seed(3)

df <- tibble(
  lon = runif(250, -97.3, -97),
  lat = runif(250, 31.4, 31.7),
  z = geoR::grf(grid = cbind(lon, lat), cov.model = "exp", cov.pars = c(3, .1), nugget = 1)$data
) 
```

Below, we set up ggmap and visualize our sample points:
```{r ggmap_setup}
library("ggmap")

theme_update(
  axis.title = element_blank(),
  axis.text = element_blank(),
  axis.ticks = element_blank() 
)

waco_map <- get_map(location = c(-97.3, 31.4, -97, 31.7), color = "bw")

ggmap(waco_map) +
  geom_point(data = df, aes(lon, lat, color = z, size = z)) +
  guides(color = guide_legend(), size = guide_legend()) +
  scale_color_viridis_c()
```

***

It is simple to include kriging layers on top of ggmap raster images, we need only adjust `alpha`:
```{r ggmap_krige_1, fig.height = 10, fig.width = 10}
p_heatmap_dots <- ggmap(waco_map) +
  geom_krige(data = df, aes(lon, lat, z = z), alpha = .4, inits = c(3, .1, 1), model = "Exp") +
  geom_point(data = df, aes(lon, lat, size = z)) +
  scale_fill_viridis_c()

p_filled_contours_dots <- ggmap(waco_map) +
  geom_krige_contour_filled(data = df, aes(lon, lat, z = z), alpha = .4, bins = 7, 
                            inits = c(3, .1, 1), model = "Exp") +
  geom_point(data = df, aes(lon, lat, size = z)) +
  scale_fill_viridis_d()

p_heatmap_dots /
p_filled_contours_dots
```

***

And now, without the plotted observations:

```{r ggmap_krige_2, fig.height = 15, fig.width = 10}
p_heatmap <- ggmap(waco_map) +
  geom_krige(data = df, aes(lon, lat, z = z), alpha = .4, inits = c(3, .1, 1), model = "Exp") +
  scale_fill_viridis_c()

p_filled_contours <- ggmap(waco_map) +
  geom_krige_contour_filled(data = df, aes(lon, lat, z = z), alpha = .4, bins = 7, 
                            inits = c(3, .1, 1), model = "Exp") +
  scale_fill_viridis_d()

p_contours <- ggmap(waco_map) +
  geom_krige_contour(data = df, aes(lon, lat, z = z), alpha = .4, bins = 7, 
                     inits = c(3, .1, 1), model = "Exp")

p_heatmap /
p_filled_contours /
p_contours
```


***
<!-- Include real world example, probably weather-based -->
<!-- We could provide similar geoms for estimating the mean of Poisson processes. -->








